Ab Initio Effective One-Electron Potential Operators. II. Applications for Exchange-Repulsion Energy in Effective Fragment Potentials

### 2019-12-30 papers/3_oep_technique.Pauli/paper-r3.tex
Revision 3 updated

\input{header.tex}
\begin{document}

@@

The exchange\hyp{}repulsion (EXR) interaction energy in the
second generation of the
effective fragment potentials (EFP2) method\cite{Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013,
   Nguyen.Pachter.Day.JCP.2014,
   Day.Jensen.Gordon.Webb.Stevens.Krauss.Garmer.Basch.Cohen.JCP.1996}
describes the non\hyp{}electrostatic repulsion
between two isolated and unperturbed Hartree\hyp{}Fock\cite{Roothaan.RevModPhys.1951} 
wavefunctions in an interacting
bi\hyp{}molecular complex.
The EXR EFP2 term was derived from the intermolecular perturbation theory
of Murrell et al.\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965,Otto.Ladik.ChemPhys.1975} 
up to the second order with respect to the wavefunction overlap,
and is given by~\cite{Jensen.JCP.1996,Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998}
%
\begin{equation} \label{e:exr-efp2}
 E^{\rm Ex-Rep} \approx
 E^{\rm Ex} +
 E^{\rm Rep}(S^{-1}) + 
 E^{\rm Rep}(S^{-2}) \;.
\end{equation}
%
In the above equation,
$E^{\rm Ex}$ is the exchange energy,\cite{Jensen.JCP.1996}
%
\begin{equation} \label{e:exc-efp2}
 E^{\rm Ex} \approx 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} 
 \sqrt{\frac{-2\ln{\vert S_{ij}\vert}}{\pi}} \frac{S^2_{ij}}{r_{ij}} \;,
\end{equation}
%
whereas the first\hyp{} and second\hyp{}order repulsion terms
are accordingly~\cite{Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998}
%
\begin{multline} \label{e:rep-efp2.s1}
 E^{\rm Rep}(S^{-1}) \approx - 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} S_{ij} \\
 \times
 \left\{
 \sum_{k\in A}^{\rm Locc} F_{ik}^A S_{kj} 
+\sum_{l\in B}^{\rm Locc} F_{jl}^A S_{li} 
 -2T_{ij}
 \right\}
\end{multline}
%
and 
%
\begin{multline} \label{e:rep-efp2.s2}
 E^{\rm Rep}(S^{-2}) \approx 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} S_{ij}^2 
 \\ \times 
 \Bigg\{
 \sum_{x\in A}^{\rm At}  \frac{-Z_x}{r_{xb}}
+\sum_{y\in B}^{\rm At}  \frac{-Z_y}{r_{ya}} 
+\sum_{k\in A}^{\rm Locc} \frac{   2}{r_{jk}}
+\sum_{l\in B}^{\rm Locc} \frac{   2}{r_{il}}
-r_{ij}^{-1}
 \Bigg\} \;.
\end{multline}
%
The indices $i$, $j$, $k$ and $l$ label the localized occupied orbitals (LMO's, denoted by `Locc')
located at their charge centroids ${\bf r}_{i(j)}$,
whereas $x$ and $y$ label atomic nuclei (denoted by `At') with atomic numbers 
$Z_{x(y)}$ located at ${\bf r}_{x(y)}$. The relative distances are
defined by $r_{uw}=\vert {\bf r}_u-{\bf r}_w\vert$.
$S_{ij}$ and $T_{ij}$ are the overlap and kinetic\hyp{}energy
integrals, respectively, whereas $F_{ik}^X$ are the Fock matrix elements
of unperturbed monomer $X$, all defined in LMO basis.
It is known that evaluating Eq.~\eqref{e:exr-efp2}
is remarkably efficient and gives usually very accurate
estimates of the reference EXR energies according to the symmetry 
adapted perturbation theory\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994} (SAPT).
The computational cost to evaluate
the exchange and second\hyp{}order terms
is much smaller than that of the first\hyp{}order term,
which requires evaluation of not only the one\hyp{}electron overlap,
but also the kinetic
energy integrals in a space of occupied molecular orbital (MO) basis.

@@

         - W^B_{ji} \\

@@

is the nucleus\hyp{}electron potential operator

@@

In our in\hyp{}house plug-in to the {\sc Psi4}
quantum chemistry
program,\cite{Psi4.JCTC.2017}
we implemented EXR energy models from Eq.~\eqref{e:exr-efp2}
(EFP2), Eq.~\eqref{e:rep.murrell-etal.S1.oep} (OEP), as well as 
the approximate intermolecular perturbation theory with exchange of 
Murrell et~al.\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965},
%from Eq.~\eqref{e:rep.murrell-etal.S1} 
the exact EXR energy of Hayes \& Stone\cite{Hayes.Stone.MolPhys.1984},
here referred to as the HS model.
The latter was
implemented in the dimer\hyp{}centered basis set\cite{Chalasinski.Gutowski.MolPhys.1985} (DCBS),
to eliminate the basis set
superposition error (DCBS) in the benchmark calculations.
%and the density decomposition scheme 
%at the Hartree\hyp{}Fock level of theory (here referred to as the DDS/HF) 
%developed by Mandado and Hermida\hyp{}Ram{\'o}n,\cite{Mandado.Hermida-Ramon.JCTC.2011} 
%for validation of the EDF technique.
%When implemented in the dimer\hyp{}centered basis set\cite{Chalasinski.Gutowski.MolPhys.1985} (DCBS),
%the DDS/HF model is
%equivalent to the exact EXR energy by Hayes \& Stone and gives identical results,
%which is the case in this work. 
%Using DCBS also allows one to eliminate the basis set
%superimposition error (DCBS).
The Boys method\cite{Boys.RevModPhys.1960} was used to localize
molecular orbitals---the same as in the original EFP2 formulation,
Note here that, since
Murrell et~al. theory is invariant with respect to unitary transformation
of molecular orbitals, the OEP\hyp{}based expression 
in Eq.~\eqref{e:rep.murrell-etal.S1.oep} is also invariant and does not require
orbital localization. 
%in which
%
%\begin{equation}\label{e:dds-exr}
% \Delta E^{\rm Ex-Rep}_{\rm DDS/HF} = 
%\end{equation}
%
%
\begin{figure*}[t]
\includegraphics[width=0.8\textwidth]{fig-2.pdf}
\caption{\label{f:fig-2} {\bf Accuracy of the OEP and EFP2 models of exchange\hyp{}repulsion energy
across various bi\hyp{}molecular systems.} 
(a) NCB31 
database,\cite{Zhao.Schultz.Truhlar.JCTC.2006,
Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005} 
of non\hyp{}covalent interactions
and
(b) BBI subset\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017} 
of backbone\hyp{}backbone interactions in proteins from the BioFragment Database.
For the OEP calculations, the EDF-1 scheme with the aug-cc-pVDZ-jkfit auxiliary basis set
was used.
} 
\end{figure*}
%

Three complexes: 
(H$_2$O)$_2$, 
H$_2$O--HOCH$_3$, and
(DMSO)$_2$,
were chosen as model systems to analyze the asymptotic dependence 
of EXR energy, which is shown in Figure~\ref{f:fig-1}. 
The reference (zero\hyp{}displacement) geometries
were obtained as described in Ref.\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019} 
and the structures  
along the translation direction are depicted
in the insets in Figure~\ref{f:fig-1}.
Energy\hyp{}optimizations were performed at the HF/6-31+G(d,p) level,
as implemented in 
the {\sc Gaussian16} quantum chemistry program package.\cite{Gaussian16}
OEP and EFP2 models correctly describe the EXR energy at all separations
for the studied model systems with very similar accuracy. 
However,
contrary to our previous application of the OEP technique\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019},
where approximately 20\hyp{}fold speedups as compared to the EFP2 model were achieved
for the evaluation of the charge\hyp{}transfer energy,
the CPU timings of EXR calculations in the OEP method are comparable but roughly 2
times higher than that of the EFP2 model
when the usual density fitting auxiliary basis sets such as aug-cc-pVDZ-jkfit or aug-cc-pVDZ-ri are used
(Table~\ref{t:tab-1}).
Computational cost of $\Delta E^{\rm Rep}(S^{-1})$ in EFP2 model
is approximately $s(on^2 + o^2n) + t(on^2 + o^2n) + 2o^3$, whereas in OEP model
is $s(on^2 + o^2n) + 2ao^2$, both of which have comparable magnitudes. 
Here, $s$ and $t$ denote the relative costs of evaluation of the overlap
and kinetic-energy one electron integrals (OEIs), whereas $o$, $n$ and $a$ is
the number of LMOs, primary AOs and auxiliary AOs, respectively.
It is clear that the critical parameter is the number of auxiliary basis functions $a$,
which should be comparable with the number of occupied orbitals $o$ 
as $t$ is usually only 2--3 times larger than $s$.
To investigate this issue more thoroughly we developed small auxiliary basis sets for H$_2$O, CH$_3$OH and DMSO, 
%developed
via the basis set optimization method described in Appendix~A to Paper I
in conjunction with the basin hopping global optimization algorithm\cite{Wales.EnergyLandscapes.2003,
Wales.Scheraga.Science.1999,
Wales.Doye.JPCA.1997,Li.Scheraga.PNAS.1987}
as implemented in the SciPy Python library\cite{SciPy.2019}
(see Supplementary Information).
%to minimize the computational cost of evaluating the summation terms
%from Eq.~\eqref{e:rep.murrell-etal.S1.oep},
Applying such basis sets
reduces computational effort appreciably and makes the OEP method on average 1.5 times faster than EFP2.
For instance, calculation of the EXR energy in the DMSO dimer by using 6-31++G(d,p) primary
basis set and 
1.2~GHz AMD EPYC\texttrademark{} 7301 16-Core Processor
requires about 4.6~ms and 3.0~ms when EFP2 and OEP models are used, respectively.
%Note that the benchmark Hayes \& Stone's calculations predict slightly larger repulsion energies

To investigate the accuracy of the OEP\hyp{}based repulsion
term across
%perform statistical error analysis for 
a variety of
interacting systems, a selection of
bi\hyp{}molecular complexes from the non\hyp{}covalent
interactions database NCB31 developed by the Truhlar's 
group,\cite{Zhao.Schultz.Truhlar.JCTC.2006,
Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
as well as the BioFragment Database subset BBI for backbone\hyp{}backbone
interactions in proteins of Sherrill group,\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017} 
as implemented in the {\sc Psi4}
program,\cite{Psi4.JCTC.2017}
were utilized. The computed
EXR energies were compared with the reference HS results (Figure~\ref{f:fig-2}).
On average, OEP and EFP2 methods tend to underestimate the EXR energy by 0.7\% and 7.6\%
in the NCB31/6-311++G(d,p) data set, and 6.0\% and 15\% in the BBI/6-31+G(d,p) data set,
respectively.
This is mostly due to BSSE, that is corrected for
only in the reference calculations with the HS model, and the effect is usually larger
when smaller primary basis sets are used.
The OEP model performs slightly better when compared to the EFP2
method for systems studied in this work, with 
root-mean-square errors of OEP and EFP2 models of
2.0 kcal/mol and 2.6 kcal/mol, respectively, for NCB31/6-311++G(d,p) data set, 
or
0.5 kcal/mol and 1.2 kcal/mol for BBI/6-31+G(d,p) data set.
The slightly better performance of OEP is most likely due to
fruitful cancellation of errors, since
the mutual differences between EFP2 and OEP estimations of 
$\Delta E^{\rm Rep}(S^{-1})$ are on the average much smaller
(RMSE between EFP2 and OEP is only 0.7\% in all data sets).
%
% RWG: Aren't they increasing with the magnitude of interaction
% energy? For instance what is the difference in magnitudes for the
% last complex in NCB31 of 48kcal/mol and the last three complexes
% in BBI above 15 kcal/mol. 
% 
% From what I see the OEP is consistently closer to the reference
% values - thus I think it is a bit more than fruitful cancellation.
% Can you find one example when EFP2 is closer to the reference?
% Have you tried to compute this in other (larger) basis sets? If
% this trend would be consistent in larger basis set then it is
% pretty obvious that it is not just fruitful cancellation.
%
Nevertheless, together with the performance data of both models
from Table~\ref{t:tab-1}, OEP model consistently outperforms the
EFP2 approach in terms of accuracy and also performance, provided
that the minimal auxiliary basis set is used.

%In particular, the subsets from the NCB31 database were separately considered:
%the HB6/04 hydrogen bonding database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
%the DI6/04 dipoleinteraction database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
%the CT7/04 charge-transfer complex database,\cite{Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
%the WI7/05 weak interaction database,\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}
%and PPS5/05 the $\pi$-$\pi$ stacking database.\cite{Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005,Zhao.Truhlar.JPCA.2005}

@@

\begin{acknowledgments}
This project is carried out under POLONEZ programme which has received funding from the European Union's
Horizon~2020 research and innovation programme under the Marie Sk{\l}odowska-Curie grant agreement 
No.~665778. This project is funded by National Science Centre, Poland 
(grant~no. 2016/23/P/ST4/01720) within the POLONEZ 3 fellowship.
\end{acknowledgments}

>>>>

\input{header.tex}
\begin{document}
%\preprint{AIP/123-OEP}

@@

The exchange\hyp{}repulsion (EXR) interaction energy 
usually describes the non\hyp{}electrostatic repulsion
between two isolated and unperturbed 
wavefunctions in an interacting
molecular 
complex.\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965,Otto.Ladik.ChemPhys.1975,Hayes.Stone.MolPhys.1984,Jeziorski.Moszynski.Szalewicz.ChemRev.1994,Jensen.Gordon.MolPhys.1996,Hesselmann.Jansen.Schutz.JCP.2005,Mandado.Hermida-Ramon.JCTC.2011}
In the
second generation of the
effective fragment potentials (EFP2) method\cite{Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013,
   Nguyen.Pachter.Day.JCP.2014,
   Day.Jensen.Gordon.Webb.Stevens.Krauss.Garmer.Basch.Cohen.JCP.1996}
the EXR energy is an important component of the total interaction energy.
It is
derived from the intermolecular perturbation theory
of Murrell et al.\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965,Otto.Ladik.ChemPhys.1975} 
up to the second order with respect to the wavefunction overlap
assuming Hartree\hyp{}Fock\cite{Roothaan.RevModPhys.1951} (HF) wavefunctions,
and is given by~\cite{Jensen.JCP.1996,Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998}
%
\begin{equation} \label{e:exr-efp2}
 E^{\rm Ex-Rep} \approx
 E^{\rm Ex} +
 E^{\rm Rep}(S^{-1}) + 
 E^{\rm Rep}(S^{-2}) \;.
\end{equation}
%
In the above equation,
$E^{\rm Ex}$ is the exchange energy,\cite{Jensen.JCP.1996}
%
\begin{equation} \label{e:exc-efp2}
 E^{\rm Ex} \approx 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} 
 \sqrt{\frac{-2\ln{\vert S_{ij}\vert}}{\pi}} \frac{S^2_{ij}}{r_{ij}} \;,
\end{equation}
%
whereas the first\hyp{} and second\hyp{}order repulsion terms
are accordingly~\cite{Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998}
%
\begin{multline} \label{e:rep-efp2.s1}
 E^{\rm Rep}(S^{-1}) \approx - 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} S_{ij} \\
 \times
 \left\{
 \sum_{k\in A}^{\rm Locc} F_{ik}^A S_{kj} 
+\sum_{l\in B}^{\rm Locc} F_{jl}^A S_{li} 
 -2T_{ij}
 \right\}
\end{multline}
%
and 
%
\begin{multline} \label{e:rep-efp2.s2}
 E^{\rm Rep}(S^{-2}) \approx 2 \sum_{i\in A}^{\rm Locc} \sum_{j\in B}^{\rm Locc} S_{ij}^2 
 \\ \times 
 \Bigg\{
 \sum_{x\in A}^{\rm At}  \frac{-Z_x}{r_{xb}}
+\sum_{y\in B}^{\rm At}  \frac{-Z_y}{r_{ya}} 
+\sum_{k\in A}^{\rm Locc} \frac{   2}{r_{jk}}
+\sum_{l\in B}^{\rm Locc} \frac{   2}{r_{il}}
-r_{ij}^{-1}
 \Bigg\} \;.
\end{multline}
%
The indices $i$, $j$, $k$ and $l$ label the localized occupied orbitals (LMO's, denoted by `Locc')
located at their charge centroids ${\bf r}_{i(j)}$,
whereas $x$ and $y$ label atomic nuclei (denoted by `At') with atomic numbers 
$Z_{x(y)}$ located at ${\bf r}_{x(y)}$. The relative distances are
defined by $r_{uw}=\vert {\bf r}_u-{\bf r}_w\vert$.
$S_{ij}$ and $T_{ij}$ are the overlap and kinetic\hyp{}energy
integrals, respectively, whereas $F_{ik}^X$ are the Fock matrix elements
of unperturbed monomer $X$, all defined in LMO basis.
It is known that evaluating Eq.~\eqref{e:exr-efp2}
is remarkably efficient and gives usually very accurate
estimates of the reference EXR energies according to the symmetry 
adapted perturbation theory\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994} (SAPT).
The computational cost to evaluate
the exchange and second\hyp{}order terms
is much smaller than that of the first\hyp{}order term,
which requires evaluation of not only the one\hyp{}electron overlap,
but also the kinetic
energy integrals in a space of occupied molecular orbital (MO) basis.

@@

         - W^B_{ij} \\

@@

is the electrostatic potential operator due to nuclei

@@

In our in\hyp{}house plug-in to the {\sc Psi4}
quantum chemistry
program,\cite{Psi4.JCTC.2017}
we implemented EXR energy models from Eq.~\eqref{e:exr-efp2}
(EFP2), Eq.~\eqref{e:rep.murrell-etal.S1.oep} (OEP), as well as 
the approximate intermolecular perturbation theory with exchange of 
Murrell et~al.\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965},
%from Eq.~\eqref{e:rep.murrell-etal.S1} 
and the exact EXR energy of Hayes \& Stone\cite{Hayes.Stone.MolPhys.1984},
here referred to as the HS model.
The latter was
implemented in the dimer\hyp{}centered basis set\cite{Chalasinski.Gutowski.MolPhys.1985} (DCBS),
to eliminate the basis set
superposition error (DCBS) in the benchmark calculations.
%and the density decomposition scheme 
%at the Hartree\hyp{}Fock level of theory (here referred to as the DDS/HF) 
%developed by Mandado and Hermida\hyp{}Ram{\'o}n,\cite{Mandado.Hermida-Ramon.JCTC.2011} 
%for validation of the EDF technique.
%When implemented in the dimer\hyp{}centered basis set\cite{Chalasinski.Gutowski.MolPhys.1985} (DCBS),
%the DDS/HF model is
%equivalent to the exact EXR energy by Hayes \& Stone and gives identical results,
%which is the case in this work. 
%Using DCBS also allows one to eliminate the basis set
%superimposition error (DCBS).
The Boys method\cite{Boys.RevModPhys.1960} was used to localize
molecular orbitals---the same as in the original EFP2 formulation.
Note here that, since
Murrell et~al. theory is invariant with respect to unitary transformation
of molecular orbitals, the OEP\hyp{}based expression 
in Eq.~\eqref{e:rep.murrell-etal.S1.oep} is also invariant and does not require
orbital localization. 
%in which
%
%\begin{equation}\label{e:dds-exr}
% \Delta E^{\rm Ex-Rep}_{\rm DDS/HF} = 
%\end{equation}
%
%
\begin{figure*}[t]
\includegraphics[width=0.8\textwidth]{fig-2.pdf}
\caption{\label{f:fig-2} {\bf Accuracy of the OEP and EFP2 models of exchange\hyp{}repulsion energy
across various bi\hyp{}molecular systems.} 
(a) NCB31 
database\cite{Zhao.Schultz.Truhlar.JCTC.2006,
Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005} 
of non\hyp{}covalent interactions
and
(b) BBI subset\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017} 
of backbone\hyp{}backbone interactions in proteins from the BioFragment Database.
For the OEP calculations, the EDF-1 scheme with the aug-cc-pVDZ-jkfit auxiliary basis set
was used.
} 
\end{figure*}
%

Three complexes: 
(H$_2$O)$_2$, 
H$_2$O--HOCH$_3$, and
(DMSO)$_2$,
were chosen as model systems to analyze the asymptotic dependence 
of EXR energy, which is shown in Figure~\ref{f:fig-1}. 
The reference (zero\hyp{}displacement) geometries
were obtained as described in Ref.\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019} 
and the structures  
along the translation direction are depicted
in the insets in Figure~\ref{f:fig-1}.
Energy\hyp{}optimizations were performed at the HF/6-31+G(d,p) level,
as implemented in 
the {\sc Gaussian16} quantum chemistry program package.\cite{Gaussian16}
OEP and EFP2 models correctly describe the EXR energy at all separations
for the studied model systems with very similar accuracy. 
However,
contrary to our previous application of the OEP technique\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019},
where approximately 20\hyp{}fold speedups as compared to the EFP2 model were achieved
for the evaluation of the charge\hyp{}transfer energy,
the CPU timings of EXR calculations in the OEP method are comparable but roughly 2
times higher than that of the EFP2 model
when the usual density fitting auxiliary basis sets such as aug-cc-pVDZ-jkfit or aug-cc-pVDZ-ri are used
(Table~\ref{t:tab-1}).
Computational cost of $E^{\rm Rep}(S^{-1})$ in EFP2 model
is approximately $s(on^2 + o^2n) + t(on^2 + o^2n) + 2o^3$, whereas in OEP model
is $s(on^2 + o^2n) + 2ao^2$, both of which have comparable magnitudes. 
Here, $s$ and $t$ denote the relative costs of evaluation of the overlap
and kinetic-energy one electron integrals (OEIs), whereas $o$, $n$ and $a$ is
the number of LMOs, primary AOs and auxiliary AOs, respectively.
It is clear that the critical parameter is the number of auxiliary basis functions $a$,
which should be comparable with the number of occupied orbitals $o$ 
as $t$ is usually only 2--3 times larger than $s$.
To investigate this issue more thoroughly we developed small auxiliary basis sets for H$_2$O, CH$_3$OH and DMSO, 
%developed
via the basis set optimization method described in Appendix~A to Paper I
in conjunction with the basin hopping global optimization algorithm\cite{Wales.EnergyLandscapes.2003,
Wales.Scheraga.Science.1999,
Wales.Doye.JPCA.1997,Li.Scheraga.PNAS.1987}
as implemented in the SciPy Python library\cite{SciPy.2019}
(see Supplementary Information).
%to minimize the computational cost of evaluating the summation terms
%from Eq.~\eqref{e:rep.murrell-etal.S1.oep},
Applying such basis sets
reduces computational effort appreciably and makes the OEP method on average 1.5 times faster than EFP2.
For instance, calculation of the EXR energy in the DMSO dimer by using 6-31++G(d,p) primary
basis set and 
1.2~GHz AMD EPYC\texttrademark{} 7301 16-Core Processor
requires about 4.6~ms and 3.0~ms when EFP2 and OEP models are used, respectively.
%Note that the benchmark Hayes \& Stone's calculations predict slightly larger repulsion energies

To investigate the accuracy of the OEP\hyp{}based repulsion
term across
%perform statistical error analysis for 
a variety of
interacting systems, a selection of
bi\hyp{}molecular complexes from the non\hyp{}covalent
interactions database NCB31 developed by the Truhlar's 
group,\cite{Zhao.Schultz.Truhlar.JCTC.2006,
Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005}
as well as the BioFragment Database subset BBI for backbone\hyp{}backbone
interactions in proteins of Sherrill group,\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017} 
as implemented in the {\sc Psi4}
program,\cite{Psi4.JCTC.2017}
was utilized. The computed
EXR energies were compared with the reference Murrell~et~al. results (Figure~\ref{f:fig-2})
as well as with the reference HS results (Figure~S1).
On average, OEP and EFP2 methods are in good agreement
with the reference models and correlation $R^2$ coefficients
are between 90--99\% in all data sets. 
The mutual differences between EFP2 and OEP estimations of 
$E^{\rm Rep}(S^{-1})$ are on the average 0.7 kcal/mol  (Figure~S2) which shows that the ERI elimination proposed here is accurate.
Root mean square errors (RMSE) of EXR energy estimation
via OEP and EFP2 models
%against both %the Murrell~et~al. and the HS reference
are around 2.0~kcal/mol and 2.6~kcal/mol in the NCB31/6-311++G(d,p) data set, and 0.6~kcal/mol 
and 0.3--1.2~kcal/mol
in the BBI/6-31+G(d,p) data set, respectively. However,
while the OEP and EFP2 models tend to overestimate
the EXR energy as compared to the Murrell~et~al. reference
by 7--12\%
(except for the EFP2 model in the BBI/6-31+G(d,p) data set that shows very good agreement), 
they consistently underestimate the EXR energy
as compared to the HS reference by 0.7--7\% (OEP)
and 8--15\% (EFP2).
This is mostly due to BSSE, that is corrected for
only in the reference calculations with the HS model.
%, and the effect is usually larger
%when smaller primary basis sets are used.
% though the.
%much smaller
%(RMSE between EFP2 and OEP is only 0.7\% in all data sets).
%with root mean square errors (RMSE)
%not larger than 2.3~kcal/mol and 2.9~kcal/mol
%in NCB31/6-311++G(d,p) data set,
%as well as
%0.6~kcal/mol and 1.2~kcal/mol 
%for BBI/6-31+G(d,p) data set. 
%On the one hand, when compared to the HS model as reference,
%OEP and EFP2 tend to underestimate the EXR energy
%in all studied cases. On the other hand, EFP2 well
%reproduces EXR energies obtained by using the Murrell~et~al. %model,
%while OEP slightly overestimates the 
%However, due to the nature of
%approximations made, the origins of errors are 
%tend to underestimate the EXR energy by 0.7\% and 7.6\%
%in the NCB31/6-311++G(d,p) data set, and 6.0\% and 15\% in the BBI/6-31+G(d,p) data set,
%respectively.
%The OEP model performs slightly better when compared to the EFP2
%method for systems studied in this work, with 
%root-mean-square errors of OEP and EFP2 models of
%2.0 kcal/mol and 2.6 kcal/mol, respectively, for NCB31/6-311++G(d,p) data set, 
%or
%0.5 kcal/mol and 1.2 kcal/mol for BBI/6-31+G(d,p) data set.
%On the other hand, when compared to 
%the Murrell~et~al. theory as reference (see Supplementary Information),
%EFP2 model performs 
%slightly better than OEP model. Note that in this case BSSE
%is not included in the reference calculations.
%Therefore
%The slightly better performance of OEP vs HS reference 
%is most likely due to
%fruitful cancellation of errors, 
%Careful inspection (Figure~S1) revealed that when compared to 
%the Murrell~et~al. theory as reference, EFP2 model performs 
%slightly better than OEP model.
% RWG: Aren't they increasing with the magnitude of interaction
% energy? For instance what is the difference in magnitudes for the
% last complex in NCB31 of 48kcal/mol and the last three complexes
% in BBI above 15 kcal/mol. 
% From what I see the OEP is consistently closer to the reference
% values - thus I think it is a bit more than fruitful cancellation.
% Can you find one example when EFP2 is closer to the reference?
% Have you tried to compute this in other (larger) basis sets? If
% this trend would be consistent in larger basis set then it is
% pretty obvious that it is not just fruitful cancellation.
Nevertheless,
OEP model is rather of roughly comparable accuracy as EFP2 model in all the systems studied.
Together with the performance data of both models
from Table~\ref{t:tab-1}, OEP model consistently outperforms the
EFP2 approach provided
that the minimal auxiliary basis set is used.

@@

\begin{acknowledgments}
This project is carried out under POLONEZ programme which has received funding from the European Union's
Horizon~2020 research and innovation programme under the Marie Sk{\l}odowska-Curie grant agreement 
No.~665778. This project is funded by National Science Centre, Poland 
(grant~no. 2016/23/P/ST4/01720) within the POLONEZ 3 fellowship.
Wroclaw Centre for Networking and Supercomputing (WCSS) is acknowledged for
computational resources.
\end{acknowledgments}
